Adopting agile software development in the large organizations: A systematic literature review

### 2014-08-09 journal/article.tex
Commented out TODO
Fixes
Preambles for article

\section{Introduction}

--> Introductory paragraph

-- Who will be interested of this?

-- What can be learned from this`?

--> Motivation for research
-- There are no systematic reviews for this
-- A collection of experience reports can be used for creating a theoretical
   base.
--> There is interest in large scale [ref XP13 workshop]

--> Freudenberg, S. and Sharp, H., "The Top 10 Burning Research Questions from
    Practitioners," IEEE Software, pp. 8-9, 2010.

--> Why is large-scale software development different from smaller scale?

--> This literature review is intended for researchers as a theoretical
   and proven foundation for the real problems that organizations face in
   transformations.
   This is also intended to be directly usable by practitioners, to serve as a
   base of general advice with strong motivation (although every organization
   has unique needs).

Dingsoyr:
The nature of software development, being innovative work that is only partially
compatible with programmed coordination, requires a strong emphasis on personal
communication [Kraut, R. E. and Streeter, L. A., "Coordination in software
development," Communications of the ACM, vol. 38, pp. 69-81, 1995]

% ---------------------------------------------------------------------
\section{Background}
\label{sec:background}

Lindvall, M., et al., "Agile Software Development in Large Organizations,"
IEEE Computer, vol. 37, pp. 26-34, 2004.

Boehm, B. and Turner, R., Balancing Agility and Discipline: A Guide for the
Perplexed: Addison-Wesley, 2003.
Boehm, B., "Get ready for agile methods, with care," IEEE Computer, vol. 35,
pp. 64 - 69, 2002.
--> Any hints on size?? Maybe size leaning towards discipline?

What is agile?
[6] Conboy K., Fitzgerald B., “Toward a conceptual framework for agile methods:
a study of agility in different disciplines,”
Proc. ACM workshop on Interdisciplinary software engineering research (WISER), 2004, pp. 37-44.

\subsection{Large scale general references from primary studies}

Laanti 2008:
[2] Kähkönen Tuomo, “Agile Methods for large organizations – building communities of practice”.
Proc. Agile Development Conf. (ADC), 2004, pp. 2-10.
[7] Eckstein Jutta, “Agile Software Development in the large: Diving into the deep”,
Dorset House Publishing, 2004. ISBN 0-932633-57-9.
[8] Leffingwell Dean, “Scaling Software Agility: Best Practices for Large Enterprises”,
Addison-Wesley Professional, 2007. ISBN-10: 0321458192.
[12] Highsmith J., Agile for the Enterprise, “From Agile Teams to Agile Organizations”,
Cutter, 2005. http://www.cutter.com/program/ fulltext/reports/2005/01/index.html .
[13] Schwaber Ken, “Enterprise and Scrum”, Microsoft Press 2007, ISBN-10: 0735623376.

Vlaanderen 2012:
22. Sidky, A.S.: A Structured Approach to Adopting Agile Practices: The Agile Adop-
tion Framework (June 2007)

Cockburn 2003:
R. Crocker, Large-Scale Agile Software Development, Addison-Wesley, 2003.

\subsection{Things that may need to be explained}

* Scrum

* XP

* Definition of sprint

* Definition of product owner

* Self organization and independence of teams -- This is a question relating how
  to organize management.

* Scrum master must not command and control -- Scrum master as manager??

* People centricity -- agile is not for squeezing efficiency -- agile is not for
  fewer career opportunities in a flatter organization.

* Team centricity -- agile focuses on team rather than individuals

* Agile focuses on short time horizons ???

* Agile is a mindset -- it can not be studied from a book -- need of coaching

* What are the agile principles? What is being talked about when ``agile
  principles'' is referred to?

* What is coaching? Why is it done?

* Principle of build often? Principle of nightly builds or CI? Is this from XP?
  Is this an agile principle?

* Kotter: Having a ``guiding team'' for change

% ---------------------------------------------------------------------
\section{Research method}
\label{sec:method}

The goal of the research was to aggregate evidence on how large organizations
transform into an agile way of working. The research was conducted as a
systematic literature review based on empirical studies.

The research process is presented starting from the research questions. After
that we present the definition of large scale which this research builds upon.
The bulk of this section concentrates on presenting the steps of the research
process, starting from the definition of inclusion criteria, continuing with
describing how the primary studies were selected, and finally presenting the
method for analyzing and synthesizing the primary studies. The section concludes
with a discussion on limitations of the method.

\subsection{Research questions}

The following research questions were set to guide the research.

\begin{itemize}

\item
RQ1: Why do large software development organizations initiate an agile
transformation?

\item
RQ2: How do large-scale agile transformations usually proceed?

\item
RQ3: What are typical challenges in a large-scale agile transformation?

\item
RQ4: What factors facilitate adopting of agile methods?

\end{itemize}

\subsection{Definition of large scale}
\label{sec:largescale}

A key question in this study was the definition of \emph{large scale}. Our goal
was to identify phenomena that appear particularly in large-scale
transformations, and which may not be present in small organizations. Large
scale requires additional coordination and makes communication difficult
compared to software development in small teams. In this section we present the
definition of large scale that is used in this study.

The definition of large scale was discussed in a workshop at XP2013
\cite{Dingsoyr2013a}. Proceeding the workshop, Dingsøyr et al.
compiled a description of the definition of large scale \cite{Dingsoyr2013b}.
The most common view in the XP2013 workshop was that the involvement of many
people is the main factor in large scale \cite{Dingsoyr2013b}.

Dingsøyr et al. conducted a brief literature search to identify what definitions
on large scale are used \cite{Dingsoyr2013b}. The results showed that large
scale had been regarded in terms of size in persons or teams, project budget,
code base size, and project duration. As examples from the results, Paasivaara
et al. \cite{Paasivaara2008} referred to an organization of 40 people and 7
teams. Berger and Beynon-Davies \cite{Berger2009} present the project cost of
over £10 million as a factor of large scale, but also highlight a team size of
over 50 people. A product with a code base size of over 5 million lines of code
was considered to represent a large scale development effort by Petersen and
Wohlin \cite{Petersen2010}. A project time of 2 years was considered as a factor
of large scale by Bjarnson et al. \cite{Bjarnason2011}, as well as a project
scope of 60-80 features.

As a part of our research, we identified further definitions of large scale. All
of these definitions referred to the number of people involved.
In early work on agile Fowler \cite{Fowler2000} regards the Crystal methodology
being suitable for up to 50 people, thus setting a boundary beyond which agile
methods would be expected to need special considerations. Also Williams and
Cockburn \cite{Williams2003} present 50 people as a limit which practitioners
and researchers has seen as the largest organization suitable for agile.
Elshamy and Elssamadisy \cite{Elshamy2006} considered a development project
large if it had a staff between 50 and 100 people, including all project
personnel. Finally, the largest numbers were presented by Moore and Spens
\cite{Moore2008} who define large scale in the scope of their research as 300
people across 3 sites.

Based on these findings we defined large scale in the scope of this research to
denote organizations with 50 or more people. These 50 people would not all need
to be developers, but they should be part of the development organization. For
instance, Scrum masters and software architects would be counted in. As some
studies presented the number of teams rather than number of people, we
correspondingly defined large scale to denote development efforts involving at
least 6 teams. Having 6 teams with an average teams size of 6 people, plus a
number of supporting staff, can reasonably be considered to form an organization
of 50 people.

% --> What will bring complexity into software? Not costs or the length of the
%     schedule. --> Having many people will bring the complexity in terms of the
%     SDM to use.
% 
% --> Boehm, B. and Turner, R., Balancing Agility and Discipline: A Guide for the
%     Perplexed: Addison-Wesley, 2003. --> There might be a 50 person boundary?
%     --> See Boehm also for list of

% --> The company must already be with N teams or N people, not growing up to the
%     required numbers. Small companies in growth are therefore excluded.

\subsection{Research process}

The research was conducted as an application of Kitchenham's
\cite{Kitchenham2007} guidelines for systematic literature review. Selection of
primary studies was done using first a keyword based database search to identify
potentially relevant sources, and then manually filtering the search result. The
filtering process was executed independently by two researchers (Dikert,
Paasivaara). The selected primary studies were coded for qualitative data
extraction by one researcher (Dikert). The results of the study were finally
elicited by aggregating and analyzing the coding of the primary documents.
Figure \ref{fig:researchprocess} presents an outline of the research process.
The entire process was audited and mentored by a third researcher (Lassenius).

\begin{figure}[b]
  \begin{center}
    \includegraphics{graphics/research_process.pdf}
    \caption{Outline of the research process}
    \label{fig:researchprocess}
  \end{center}
\end{figure}

We deviated from Kitchenham's guideline on the part of data extraction. Instead
of using data extraction forms we analyzed the primary studies by coding. This
deviation was made in order to minimize how possible preconceptions would affect
the data extraction step, and to use a specialized software tool that allowed us
to record evidence in as much detail as possible.
% --> Possibility to find many different viewpoinits for the research questions
% --> Lots of items, which would have lead to much diversity in data extraction form

\subsubsection{Inclusion criteria}
\label{sec:inclusioncriteria}

Based on the research questions and intended focus of the research, we defined
four facets to guide inclusion/exclusion decisions: agile software engineering,
focus on transformation, large scale, and empiricality. The first facet covers
primary studies focusing on software engineering organizations applying or
striving to apply agile methods. The second facet states that included studies
must provide insights relevant to organizational transformation, and
specifically to the research questions. The third facet underlines the
particular contribution we aim to provide with this research, namely the
exclusive focus on large organizations. The fourth facet limits inclusion to
studies presenting real world cases. These four facets were used for inclusion
and exclusion in each study selection step, including search string design,
filtering by abstracts, and full text filtering.

Table \ref{table:facets} lists the facets and gives examples on matching topics
and not relevant topics. For a study to be included it needed to be relevant
in all facets. Even if a study discussed some of the irrelevant topics it could
be included as long as it had some content matching all the facets.

\begin{table}
    \centering
    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.20\textwidth}
                     >{\raggedright\arraybackslash}p{0.34\textwidth}
                     >{\raggedright\arraybackslash}p{0.33\textwidth} }
        \toprule
        Facet  &  Relevant topics  &  Not relevant topics  \\
        \midrule
        Agile software development  & 
                The organization develops software;
                The development method introduced is agile  &

                Agile manufacturing;
                Scrum in management boards  \\

        Organizational transformation  &
                Presenting insights in the process of transformation  &

                Comparison of before and after;
                How agile is being used in large scale   \\

        Large scale    &
                At least 50 development personnel or 6 teams uses agile &

                Scaling up from small;
                A single agile team in a large setting \\

        Empiricality   &
                Case studies, experience reports  &

                Textbooks, student experiments, theory papers  \\
        \bottomrule
    \end{tabular}
    \caption{Facets of inclusion criteria with examples of relevant and
             not relevant topics}
    \label{table:facets}
\end{table}

Examples on topics excluded by the facet of \emph{agile software development}
are agile manufacturing and applying Scrum practices in management boards, as
these do not relate to software engineering. In addition, we required that the
organizational transformation was aimed to introduce agile methods, which
excluded other development methods than agile \cite{Sagesser2013}, and use of
agile methods in other contexts than software development \cite{Hodgkins2007}.

The facet of \emph{organizational transformation} was interpreted so that the
primary study must present insights on the process of transformation. This
was the most important facet, as the pivot of this research is to examine
how the transformations happen. Examples on excluded topics are comparing the
original and agile development methods \cite{Petersen2010}, discussing use of
agile in a large organization but not describing how the new methods were
introduced \cite{Mishra2011}, and merely presenting agile tools in large-scale
use \cite{Kim2012}. These sources were excluded as they did not provide insights
on organizational transformation.
% Also [Fitzgerald 2005] on just presenting use of agile methods, no introducing

Transformation and ``scaling up'' of agile practices in use are very closely
related concepts, and in some cases they are one and the same. For instance, if
transformation begins with a pilot and spreads gradually through the
organization, the process can very well be seen as a ``scaling up'' journey.
However, we excluded cases where an initially small organization was scaled up
\cite{Maranzato2012}, and discussions focusing on processes or tools without
describing organizational change \cite{Lyon2008}.
% Also [Greening 2010] on process centered viewpoint without transformation

The facet of \emph{large scale} was interpreted as presented in section
\ref{sec:largescale}, having 50 development personnel or 6 teams as as the limit
for large scale. In some cases the source presented very vague indicators on
size. For instance, we deemed the case presented by Cloke \cite{Cloke2007} to be
included as there was indications of large-scale considerations, although the
size remained unclear. The case presented by Miller and Haddad \cite{Miller2012}
was deemed to be excluded as there was no indications whatsoever on size. A
borderline exclude was the case presented by Tudor \cite{Tudor2006} as there was
no indication of issues indicating large size, although there were several
development teams.
To complicate matters, some papers talked about ``the team'' in singular when
referring to the organization \cite{Hodgkins2007}, making it clearly nontrivial
to judge whether an organization was large based on the choice of words of the
author.

Further examples on exclusion by the \emph{large scale} facet were cases with
large organizations but only a single team adopting agile \cite{Fulgham2011}.
We considered cases of single teams (although in large organizations) unrelated
to this research as our focus is on transformation of the entire organization.
Also piloting cases that resulted only in single teams using agile were excluded
\cite{Scott2008}. Finally, cases where the organization was growing to large
scale, but did not meet the size criteria at the start of the transformation,
were excluded.

% Also [Jensen 2003] with only a single team
% Possibly [Mc Cormic 2005] on single team piloting

According to the facet of \emph{empiricality} we excluded studies that did not
discuss a distinguishable real world case. We excluded textbooks, studies merely
presenting theories, and other studies that did not include any case
organization. We also excluded studies of benefits or limitations of agile in
general. Lastly, we excluded student experiments as it is implausible to
simulate the dynamics of a large organization in an experiment.

\subsubsection{Preliminary searches}

Before proceeding with identifying the primary studies a number of preliminary
searches were performed. The purpose of the preliminary searches was to create a
benchmark of potentially relevant papers that should be matched by the actual
search, and to evaluate different search strings. We started by examining top
ranked hits by trivial keywords that the more complex final search string might
miss. Initial searches were made using keywords that were as general as
possible, including ``agile transformation'' and ``large scale agile''.
Secondly, a trial keyword search was done in the selected databases (listed in
table \ref{table:databases}) using the following search string:

\begin{verbatim}
((agil* OR scrum OR xp OR lean) AND
(transform* OR transit* OR change OR migrat*) AND
((large AND (scale OR organization))) OR enterprise)
\end{verbatim}

From these preliminary searches we compiled a list of 117 possibly relevant
papers that the actual database search would be benchmarked against.

\subsubsection{Identification of primary studies}

The gathering of potential primary studies was based on a search in online
databases. The databases searched are listed in Table \ref{table:databases}. All
databases supported use of complex Boolean logic in searches, which allowed us
flexibility in constructing the search strings. The main task in the database
search was to construct a search string which yielded a reasonably large result
set. The set of matched primary studies was later amended with a small number
of studies found in references of the matched papers.

\begin{table}
    \centering
    \begin{tabular}{ l l l }
        \toprule
        Database         & URL                         & N of matches   \\
        \midrule
        IEEExplore       & http://ieeexplore.ieee.org      & 745 \\ 
        ACM              & http://dl.acm.org               & 168 \\
        Scopus           & http://www.scopus.com/home.url  & 1596 \\
        Web of Knowledge & http://apps.webofknowledge.com  & 786 \\
        \bottomrule
    \end{tabular}
    \caption{Databases included in search, and number of matched articles}
    \label{table:databases}
\end{table}

We proceeded with constructing a search string based on the facets presented in
section \ref{sec:inclusioncriteria}. However, preliminary searches had showed
that picking keywords with good precision was difficult. Especially the facet
\emph{large scale} was difficult to represent with precise words. Also the facet
of \emph{empiricality} could not sensibly be represented by keywords. We decided
to include only the facets \emph{agile software development} and
\emph{organizational transformation} in the keyword search, with the consequence
of increasing the manual filtering effort in the subsequent steps.

Having a substantial part of the matches in non-relevant areas such as agile
manufacturing, we included only articles including the term ``software'' or
articles published in relevant conferences. Instead of engineering the search
terms any further, which proved to result in excluding some relevant articles,
we chose to rely more on manual filtering. The resulting facets and keywords are
listed in Table \ref{table:searchterms}.

\begin{table}
    \centering
    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.22\textwidth}
                     >{\raggedright\arraybackslash}p{0.70\textwidth} }
        \toprule
        Facet                  & Keywords   \\
        \midrule
        Agile methods (before \& after) &
            \texttt{agile, scrum, "extreme programming",
            waterfall, "plan-driven", RUP}\\
        Organizational transformation &
            \texttt{transform*, transiti*, migrat*, journey, adopt*, deploy, introduc*,
            "roll-out", rollout}\\
        Only software related articles &
            \texttt{(software OR (conference="agile, xp, icgse, icse"))
            AND NOT (title+abs="manufacturing" OR conference="agile manufacturing")}
        \\
        \bottomrule
    \end{tabular}
    \caption{Facets and related search terms used}
    \label{table:searchterms}
\end{table}

The database keyword search matched 1875 unique papers. The results of the
search were gathered for the selection step using reference management software.

\subsubsection{Study selection}

The set of potential primary studies identified by the database search was
refined in two steps, by first filtering out abstracts and finally full text
filtering. The study selection process starting from the database search is
outlined in figure \ref{fig:selectionprocess_full}. The result of the database
search was amended with some results from the preliminary search and relevant
references from all examined papers.

\begin{figure}
  \begin{center}
    \includegraphics[width=1\textwidth]{graphics/research_process_full.pdf}
    \caption{The study selection process}
    \label{fig:selectionprocess_full}
  \end{center}
\end{figure}

The database keyword search had matched 1875 unique papers. The abstracts of
these papers were categorized independently by both Dikert and Paasivaara into
three categories: include, exclude, and uncertain. There were 1578 exclusions
and 62 inclusions that both researchers agreed on. The inclusion decisions
for the 235 abstracts with disagreement or uncertainty were resolved through
discussion. At this stage papers were excluded only if both researchers deemed
it clearly irrelevant, including any uncertain cases for full text filtering. As
a result 170 papers were selected for full text filtering.

% Inter-rater statistics
%
%                     Maria
%            excl   ???   incl    TOT
%      excl  1578    18    15    1611
% Kimi ???     84    17     8     109
%      incl    51    42    62     155
%      TOT   1713    77    85    1875
%
% ???'s =  169
% in/ex = 1706
%
% Pr(Kimi, in) = in / (tot)  =  155 / 1857 = 0.0619
% Pr(Kimi, ex) = ex / (tot)  = 1611 / 1857 = 0.8675
% Pr(Maria, in) = in / (tot) =   85 / 1857 = 0.0458
% Pr(Maria, ex) = ex / (tot) = 1713 / 1857 = 0.9225
%
% Pr(e) = Pr(Kimi, in) * Pr(Maria, in) + Pr(Kimi, ex) * Pr(Maria, ex) = 0.8031
%
% Pr(a) = (ex-ex + in-in) / tot = 0.8831
%
% k = 0.4063

Full text filtering was performed by evaluating each article against the four
facets of the inclusion criteria. Filtering was done in two steps. In the first
step Dikert extracted data relevant to the four facets. Based on the extracted
data 76 papers could be immediately deemed as included or excluded. The
remaining 94 papers were evaluated against the inclusion criteria by both Dikert
and Paasivaara, and a decision was made after discussing each paper separately.
In difficult cases Lassenius was consulted to reach a decision. As a result of
the full text filtering 47 papers were selected to be included.

We evaluated the result of the database keyword search against the benchmark
created in the preliminary search step. We concluded that 75 of the 117
preliminarily selected papers were matched by the database keyword search.
The missed 32 preliminary papers were examined, resulting in including 3
additional papers as primary sources.

In parallel with the full text filtering step the references of all papers were
also examined for relevance. Most papers used references very scarcely, typically
referencing well known descriptions of agile methods. We included 2 referenced
papers in full text analysis.

Finally, counting 47 papers from the full text filtering step, 3 papers from
preliminary searches, and 2 papers from references, 52 papers were selected
as primary studies.

\subsubsection{Handling of duplicate reports on a single case}

In several cases there were more than one primary study presenting the
transformation of the same organization. Duplicate descriptions of the same
organization focused typically on different aspects. For example, one paper
would highlight the viewpoint of developers \cite{Fry2007}, and another would
consider the transformation from user experience designers' point of view
\cite{Federoff2009}.

Even if the transformation of a single organization was described in many
studies, all sources that passed the inclusion criteria were included. Studies
presenting the same organization were treated as one unit so that we could gain
as much insight on each organization as possible. Conversely, there were also a
few papers that presented multiple case studies, and in those cases we treated
each studied organization individually.

As a result 42 unique organizations were identified in the primary studies. We
use the term \emph{study} to refer to the primary study publications, and the
term \emph{case} to refer to an individual case organization that may be
described in several different studies.

\subsubsection{Study quality assessment}

The primary research for this literature review consists almost exclusively of
industry experience reports. There were only 6 case studies with a research
method, and observations on transformation were presented only as a minor part
in these studies. Based on this finding we concluded that case studies
presenting viewpoints on organizational transformations are very scarce in
software industry. We deemed that the results would be distorted heavily and
many valuable studies would be left out if a strict quality assessment would be
part of the inclusion criteria. As a result we decided to include all experience
reports, regardless of the perceived objectivity.

The absence of research methods and subjective viewpoints of experience reports
can be seen as a factor lessening the evidence. Because of this we were
refrained from making quantitative interpretations of the data, but based the
results around describing qualitative observations. Regardless of the bias of
the primary studies, we think that the concepts that emerged from the source
material should not be questioned on a fundamental level. The overall existence
of organizational phenomena should not be questioned based on possible author
bias in experience reports.

% --> Large organizational changes are hard to measure experimentally (controlled
%     experiments?). Also due to this any student research was excluded.

% --> It may still be interesting to make some quantitative observations such as
%     a particular problem being reported in the majority of reports.

% * The primary studies did not make distinction between method, results, and
%   discussion
% 
% * The studies neither did any implications for generalizability, but rather
%   presented narratives or experiences. Some cases presented lists of
% %   recommendations.

As presented in the previous section, we allowed several descriptions of the
same case to be included. Instead of using the most complete paper as suggested
by Kitchenham \cite{Kitchenham2007}, we combined the results presented in each
paper and considered the case as a single unit. Keeping in mind the potential
bias caused by duplicate publications, including all papers enabled us to have a
more detailed view of the case organization.

\subsubsection{Coding of primary studies}
\label{sec:coding}

The primary studies were coded using the Atlas TI qualitative data analysis software.
An integrated deductive and inductive approach was chosen for coding the primary
studies \cite{Cruzes2011a}. The coding was designed to have a contextual part
and a findings part as also suggested by Cruzes and Dybå \cite{Cruzes2011a}. A
list of codes was established for contextual information, as presented in Table
\ref{table:contextualcodes}. Codes related to the research questions were chosen
to be created by an inductive process. The reason for inducting codes rather
than using an a priori approach was to avoid the researchers' previous
assumptions of the research area to affect the choice of codes

\begin{table}[h]
    \centering
    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.26\textwidth}
                     >{\raggedright\arraybackslash}p{0.66\textwidth} }
        \toprule
        Contextual code     & Explanation   \\
        \midrule

Agile method & Agile methods used in the organization (eg. Scrum, XP) \\

Business area & The business area in which the organization operates. \\

Organization size & Mentioning the size of the case organization. \\

Time of transformation & When the transformation has been in
progress, or how long the transformation has taken. Possibly relative to
the time of the research. \\

Research process & The paper describes a research process. \\

Geographical location & Where the organization is located. \\

Large scale definition & A definition of large-scale software development. \\

Multisite / GSD & Mentioning of a multisite organization. \\

        \bottomrule
    \end{tabular}
    \caption{Context information}
    \label{table:contextualcodes}
\end{table}

To give direction for code induction seven different code families were
established on beforehand, as presented in Table \ref{table:codefamilies}. A
description defining scope and examples for codes were defined for each family.
The example codes would not necessarily be used as actual codes, and would not
represent the entire or final scope of the families.

\begin{table}[t]
    \centering
    \begin{tabular}{ >{\raggedright\arraybackslash}p{0.31\textwidth}
                     >{\raggedright\arraybackslash}p{0.61\textwidth} }
        \toprule
        Code family          &  Description
        \\
        \midrule

        RQ1: Reason to change &
        Reasons to start the transformation (demand for faster delivery). \\

		RQ2: Transformation process &
		Statements that describe the transformation process (top-down, big bang,
		step wise). \\

		RQ3: Challenges &
		Statements that present challenges in the transformation (change resistance).
		\\

		RQ3: Success factors &
		Statements that present success factors in the transformation (management
		support).
		\\

		Investing in change  &
		Factors that present how the organization is investing in the
		transformation (training, consultants, tools). \\

		Practices &
		Distinguishable practices that are used or established during transformation.
		Factors that are presented as neutral in terms of successes and challenges
		in the primary studies.
		(coaching, piloting, continuous integration, communities of practice) \\

		Contextual &
		Contextual codes defined in Table \ref{table:contextualcodes}. \\

        \bottomrule
    \end{tabular}
    \caption{Code families}
    \label{table:codefamilies}
\end{table}

The inductive coding was done according to the guidelines described below.

A passage of text that presents any concept relating to the research questions
or the contextual codes becomes a quotation. If the quotation corresponds
closely to a concept that has already been coded the existing code is used.
Otherwise a new code is created, and the code is assigned to one of the
predefined code families. The labels of existing codes may need to be adjusted
when new quotations are assigned. A clarifying description is written for every
code, so that the precise meaning of the codes can be reviewed. Many codes may
be assigned to one quotation and quotations of different codes may be
overlapping.

The length of the text passage that make up a quotation may vary. For the
contextual codes the quoted passage should be of minimal length. For other
quotes the length of the quotation should be long enough to highlight the
context of the concept the quotation, varying from a few sentences to one
paragraph, or even several paragraphs. As an example of context of the
quotation, if a transformation challenge is described by a quotation it would be
good if also the source or resolution of the challenge could be included in the
quoted text passage. Having enough context should broaden the understanding
of the coded concept when looking at the quotation in isolation.

% -- Considering that mentioning the inclusion of a factor as a success, then
%    would the absence of the same factor be a challenge?

The total number of codes and quotations created in the coding process is
presented in Table \ref{table:codecount}. The total number of quotations is less
than the sum of quotations in the categories because a single quotation may have
multiple codes.

\begin{table}[t]
    \centering
    \begin{tabular}{ l r r }
        \toprule
        Code family    &  Codes  &  Quotations
        \\
        \midrule
        RQ1: Reason to change &        30 &  123 \\
		RQ2: Transformation process &  16 &  580 \\
		RQ3: Challenges &              40 &  323 \\
		RQ3: Success factors &         44 &  260 \\
		Investing in change  &          5 &  137 \\
		Practices &                    11 &  170 \\
		Contextual &                    8 &  215 \\
		Total &                       154 & 1575 \\
        \bottomrule
    \end{tabular}
    \caption{Number of codes and quotations per code family}
    \label{table:codecount}
\end{table}

\subsubsection{Validation of coding}

Coding was done by only one researcher due to resource constraints and volume of
source material. As recommended by Kitchenham \cite{Kitchenham2007} we conducted
an experiment with the goal to assess the consistency of coding that two
independent researchers may achieve.
In the experiment Dikert and Paasivaara coded independently a sample of five
primary studies. The sample was hand picked as a representative subset of the
source material. The coding was done relying solely on the guidelines presented
in the previous chapter, and the researchers were careful not to discuss the
choice of codes or interpretations relating to codes before the experiment was
completed. Finally the independent codings of the two researchers were analyzed
with two metrics: by the number of matching themes found, and by the amount of
overlap in quotations similar codes have.
% See Kitcenham, chapter 6.4

The metric of matching themes measures how well the independent researchers are
able to induct similar codes from the source material. As codes are inducted it
is likely that two researchers label similar quotes differently. To overcome
this problem we defined higher level themes under which the individual codes
were organized. Through discussion we made decisions on whether differently
labeled codes could be organized under the same theme or not. Table
\ref{table:codingexperiment} presents the total number of themes the researchers
found in each document, and how many of the themes were matching. We concluded
that on average two of three themes found by one researcher were similarly found
by the other researcher, but one of three themes was not found by the other
researcher.

The metric of quotation overlap measures how well the quotations for similar
codes match between two independent researchers. The intent is to measure how
well codes that are deemed similar by their names match in the actual quoted
text. We assigned a quotation overlap score for each code in themes that both
researchers had found. A score of 3 was given for matching codes whose quotation
content was virtually the same for both researchers. A score of 2 was given if
the quotations were mostly the same, but other researcher had included one text
passage that the other had missed. A score of 1 was given if roughly half of the
quotation was the same. If the quotations had less similarity than this a score
of 0 was given. One code provided typically 1 to 3 quotations per document. The
quotation overlap score was counted including only codes that had matching
themes. We concluded that the average quotation overlap score was 2.2.

\begin{table}
    \centering
    \begin{tabular}{ l r r r }
        \toprule
        Document    &  Matching themes  &  Total themes  &  Quotation overlap (0-3) \\
        \midrule
        P1          &  14  &  19   &  2.1  \\
        P2          &   6  &  11   &  2.4  \\
        P3          &   7  &  11   &  2.1  \\
        P4          &   2  &   4   &  2.3  \\
        P5          &  12  &  16   &  2.0  \\
        \bottomrule
    \end{tabular}
    \caption{Result of independent coding experiment}
    \label{table:codingexperiment}
\end{table}

Combining the quotation overlap result and the result that inductive coding
yielded similar themes with a factor of two thirds, we conclude that two
different researchers should agree completely on at least half of the quotations
in the source material of our study. This agreement includes both that the
researchers select the same text passage as a quotation, and that researchers
agree with the meaning of the quotation. Further interpretation of the
quotations and codes will still remain a subjective matter.

\subsubsection{Synthesis of primary studies based on coding}

The synthesis of findings was created based on the result of coding presented in
section \ref{sec:coding}. An initial organization of codes into categories was
made based on the labels of the codes. Each codes was linked to a single
research question, so the categorization of codes was made separately within
the topic of each research question. The categories of codes were labeled
appropriately to give a high level description of the linked codes. Some codes
having minor importance and few quotations were left out.

After assigning each code to a unique category, the content of the categories
was studied. Each category was studied by reading through each quotation of each
code included. Typically the quotations were displayed and examined in their
original context, considering also surrounding paragraphs of the quotation.
Notes were made on each quotation presenting noteworthy observations. Based on
the notes, an accurate description emerged for each code.

As an accurate description had been created for each code, the initially defined
categories could be refined. Some codes were recategorized, and the definitions
of a few categories were revisited, until a final ordering was reached. The
results are presented according to the final categorization.

% * If a code had more quotations there were more observations made (longer
%   result text)

\subsection{Threats to validity and limitations}

To assess the validity of this research internal validity and external validity
needs to be considered. In this section we present the identified threats to
validity and the strategies used to mitigate them. We also highlight two
specific limitations in the research: weakness in keyword search, and the
strength of evidence of primary sources.

The threat to internal validity is subjective bias, which affects two parts of
the research. First, the selection of the primary studies may have been
distorted by interpreting the inclusion criteria falsely. This risk was
mitigated at start by using three researchers in designing the inclusion
criteria. When the inclusion criteria was subsequently applied the initial
filtering was performed independently by two researchers, and unclear cases were
resolved by case-by-case discussions between two or three researchers.

The second part of the research that may have been affected by subjective bias
was the elicitation of results by coding and analysis. Our tools for mitigating
this threat were limited, as the work stages in question are particularly
laborious. For resource constraints the process could not be duplicated. We
assessed the bias of the coding step using an experiment, and prevented
subjective bias in analysis by making the results as traceable as possible. The
coding experiment revealed that two researchers can find roughly 2 of 3 themes
in an experience report identically. Traceability of results was achieved by
supplying references to each claim presented in the results.

This study had a minor threat to external validity. The systematic literature
review approach includes steps to identify all studies that may be relevant in
the study area. As long as the systematic approach is adhered to, including
systematic identification of research, threats against external validity are
minimized. External validity was also confirmed by saturation of the primary
studies. Saturation was evident, as when the majority of the primary studies had
been coded almost all quotations for the remaining studies could be assigned to
existing codes.

A particular problem in this systematic literature review was the limitations of
Boolean keyword searches in online databases. A keyword search can not identify
the facet of \emph{large scale}, and also \emph{transformation} is difficult to
capture with keywords. Because of this we assume that it is likely that a some
studies have been missed by the keyword search, potentially causing bias. To
eliminate the potential bias we would suggest manually searching the entire
archive of \emph{Agile Conference} and \emph{Agile Processes in Software
Engineering and Extreme Programming} for relevant studies. These two conferences
provided more than two thirds of the primary sources.

Another particular problem was that few of the primary studies presented a
research method. Most of the studies were experience reports where the author
was a member of the organization discussed. Because of this we must assume that
the primary studies are biased to some extent. Due to this risk of bias we chose
not to make quantitative interpretations in the results, and use only
qualitative analysis. We conclude that in the current state of research it is
necessary to include studies with varying strength of evidence in order to
relevantly aggregate evidence on large scale transformations.

% ---------------------------------------------------------------------
\clearpage

\section{Results}
\label{sec:results}

This section presents the findings of the literature review. In the first
subsection a meta analysis of the analyzed papers is presented. In the
subsequent sections findings are presented organized according to the research
questions.

The methods used for extracting the results were described in section
\ref{sec:method}. Note that the term \emph{study} is used when referring to a
primary study publication, and the term \emph{case} is used when referring to an
organization which may be described in several studies.

\subsection{Overview of primary studies}

The results of this study include findings from 52 primary studies presenting
how 42 different large software organizations introduce agile methods. Most of
the included studies were experience reports (45 studies), and in many cases it
was evident that the author had been personally involved in the transformation.
Only 6 of the included studies had a research method explicitly stated. One of
the studies was an interview article. The publication forums of the primary
studies were distributed so that 57 sources were conference proceedings, 4
sources were journal articles, and one source was a technical report.

Figure \ref{fig:transformation_time} summarizes the publication years of the
primary studies, and the years when the transformations were reported to have
started. All studies and transformations were dated after year 2000. There were
peaks in transformation studies in years 2008 and 2011. The studies were
typically published two years after the start of the transformation.

\begin{figure}[b]
  \begin{center}
    \includegraphics[width=1\textwidth]{graphics/transformation_time.pdf}
    \caption{Publication years and reported start years of transformations}
    \label{fig:transformation_time}
  \end{center}
\end{figure}

The size of the organizations varied from the minimum included size of 50 to
18,000 people. The median size was 300 people. In 7 studies the size was
presented in terms of teams, ranging from the minimum of 6 teams to 150 teams.
The median was 10 teams. In 8 studies there was no direct indication on size
but the issues discussed revealed that the organization was of large size
according to our definition. The distribution of organization sizes is presented
in figure \ref{fig:organization_size}.

\begin{figure}[!t]
  \begin{center}
    \includegraphics[width=1\textwidth]{graphics/organization_size.pdf}
    \caption{Distribution of organization sizes in primary studies}
    \label{fig:organization_size}
  \end{center}
\end{figure}

\begin{table}[!t]
    \centering
    \begin{tabular}{ l r }
        \toprule
        Business area    &  Case organizations   \\
        \midrule
		Online services for consumers and business  &  9  \\
		Telecommunications                          &  7  \\
		Enterprise management                       &  5  \\
		Banking and financial services              &  4  \\
		Health care                                  &  3  \\
		IT services                                 &  2  \\
		Government                                  &  1  \\
		Information security                        &  1  \\
		Not specified                               & 10  \\
        \midrule
		Total                                       & 42  \\
        \bottomrule
    \end{tabular}
    \caption{Business areas of case organizations in primary studies}
    \label{table:businessareas}
\end{table}

>>>>

\input{content.tex}
